---
title: "HW 16 Problem 22"
author: "Erick Castillo"
date: "4/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
tabs <- read.csv("~/M440B_HW/M440B Stuff/Data/tablets2.csv")
tabs
```

\textbf{22.} The table for this question has 7 treatments with 10 samples in each. The following is the output for the code available on Canvas.
```{r}
t1 <- stack(tabs)
t2 <- aov(values~ind,data=t1)

summary(t2)

TukeyHSD(t2)
```

The above output suggest the following:\
\
The Analysis of Variance (AOV) table indicates that there is very strong evidence to suggest that one of the differential effects $(\alpha_i)$ for $i = \{1,...,7\}$ is not equal to the others.\
\
The Tukey function output indicates that the means of the following treatments are significantly different:
\begin{itemize}
\item Lab 1 with Lab 5.
\item Lab 2 with Lab 4.
\item Lab 2 with Lab 5.
\item Lab 2 with Lab 6.
\item Lab 2 with Lab 7.
\item Lab3 with Lab 5.
\end{itemize}

The function also indicates that the following treatments have weak evidence that their means are different:
\begin{enumerate}
\item Lab 1 and Lab 4.
\item Lab3 and Lab 4.
\item Lab 5 and Lab 7.
\end{enumerate}