---
title: "M440B HW 17"
author: "Erick Castillo"
date: "5/8/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\textbf{Problem 28:} I will be using the watchstacked.csv file available on Canvas for this problem.
```{r}
watch <- read.csv("~/M440B_HW/M440B Stuff/Data/watchstacked.csv")
watch.aov <- aov(WatchData~Group,data = watch)
summary(watch.aov)
```

The above ouput is a parametric ANOVA table. Because the p-value is so high, we conclude that there is insufficient evidence to conclude that the different watches have different means.
```{r}
kruskal.test(WatchData~Group, data=watch)
```

Above is the output for the non-parametric analysis. Notice that the p-value is still big, so the conclusion would be the same; however, it is interesting to see that the p-value for the Kruskal-Wallis test is more "significant" than the parametric analysis.\
\
\textbf{Problem 27:} I will be using the mice.csv file for this problem.
```{r}
mice <- read.csv("~/M440B_HW/M440B Stuff/Data/mice.CSV")
pairwise.t.test(mice$Count, mice$Species, p.adj = "bonf")
```

The above output indicates that there is a significant difference in the mean "aggressiveness" of each of the recorded species of mice.