---
title: "M440B HW 11"
author: "Erick Castillo"
date: "3/31/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\textbf{Problem 1:} This problem was done on a separate piece of paper that will be submitted alongside this assignment.\
\
\textbf{Problem 2:} I begin this problem by inputting and plotting the given data.
```{r, out.width='50%', out.height='50%',fig.align='center'}
y = c(0.27,1.34,-0.53,0.35,1.28,-0.98,-0.72,-0.81,0.64,-1.59)
x = c(0.34,1.38,-0.65,0.68,1.40,-0.88,-0.30,-1.18,0.50,-1.75)
plot(x,y)
```

\textbf{A.} Fit a line $y=a+bx$ by method of LS and display the line on the plot.
```{r,out.width='50%', out.height='50%',fig.align='center'}
mod1 = lm(y~x)
summary(mod1)
plot(x,y)
abline(mod1)
```

The above summary table output indicates that $a=-0.0334$ and $b=0.90441$.\
\
\textbf{B.} Fit a line $x=c+dy$ by method of LS and display the line on the plot.
```{r,out.width='50%', out.height='50%',fig.align='center'}
mod2 = lm(x~y)
summary(mod2)
plot(y,x)
abline(mod2)
```

The above output indicates that $c=0.03313$ and $d=1.05501$. In this case, y is being used to predict values of x.\
\
\textbf{C.} Are the lines the same? No they are not. This is to be expected because the variables were swapped in parts A and B. That is, x was a predictor for y in part A, while y was a predictor for x in part B.\
\
\textbf{Problem 15:} This problem was done on a separate piece of paper that will be submitted alongside this assignment.\