---
title: "M440B HW 5"
author: "Erick Castillo"
date: "2/18/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\textbf{Permutation t-tests}\

\textbf{Problem 1}\

a. A permutation test makes less of an assumption about the data than a t-test. Where a t-test assumes the data is normally distributed, the two sample permutation test only assumes the data comes from independent i.i.d samples.

b. There would be a total of ${130 \choose 65}$ total permutations. This value is $\approx9.5067 \times 10^{37}$.

c. Execute and understand the provided code:
```{r, echo = FALSE, out.width='50%',out.height='50%'}
bodtmp = bodytemp <- read.csv("~/M440B_HW/bodytemp.csv")
attach(bodtmp)

#ordinary 2 sample t-test.
#t.test(temperature~gender, var.equal=TRUE)

mean.diff<-mean(temperature[gender==1])-mean(temperature[gender==2])

#the following function generates resamples of the data from gender 1 and 2.
one.test <- function(x,y) {
  xstar<-sample(x)
  mean(y[xstar==1])-mean(y[xstar==2])
  } 

#the code below uses the function defined above 1000 times.
many.meandiffs <- replicate(1000, one.test(gender,temperature)) 

#this code will generate the permutation distribution
hist(many.meandiffs)
abline(v=mean.diff, lwd=2, col="purple")
```
d. The figure indicates that there is a difference between the temperatures of females and females.

e. Calculate the above p-value. It is below:
```{r, echo=FALSE}
lpval = findInterval(mean.diff, sort(many.meandiffs))/1000
lpval
```
f. The strength of the evidence based on the approximate two sided p-value:
```{r, echo=FALSE}
rhs = findInterval(-1*mean.diff, sort(many.meandiffs))
rpval = (1000-rhs)/1000
twopval = (lpval+rpval)
twopval
```
The above p-value is significant at $5\%$ thus there is sufficient evidence to suggest that the mean body temperatures are not equal.

\newpage

\textbf{Problem 2}\
For each gender perform a one-sample permutation t-test of the hypothesis that the mean body temperature is 98.6 degrees.

\textbf{Response:}\
The following answers were gathered utilizing the table provided:\
\
\underline{Males}: the one sided p-value is $\frac{5}{32} \approx 0.156$ and the two sided p-value is $\frac{10}{32} \approx 0.313$. Both p-values are not significant, thus in both cases I would fail to reject the null hypothesis; the null hypothesis being $H_0: \mu = 98.6$.\
The one sided mid p-value for males is $\frac{4.5}{32} \approx 0.141$ while the two sided mid p-value is $\frac{9}{32} \approx 0.281$. Once again both these p-values are not significant, thus I would fail to reject the null hypothesis.\
\
\underline{Females}: the one sided p-value is $\frac{7}{32} \approx 0.219$ while the two sided p-value is $\frac{14}{32} \approx 0.438$. Both p-values are not significant, thus I would fail to reject the null hypothesis.\
The one sided mid p-value is $\frac{6}{32} \approx 0.188$ and the two sided mid p-value is $\frac{12}{32} = 0.375$. Once again the p-values are not significant, thus leading once again to a failure to reject the null hypothesis.\
\
For both sexes, the samples drawn did not provide sufficient evidence to show that the average body temperature is much different from 98.6 degrees.




