---
title: "M440B HW 9"
author: "Erick Castillo"
date: "3/15/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
The following is a function I will use throughout the HW when performing $\chi^2$ tests.
```{r}
pears_chi <- function(obs, exp){
  return(sum((obs - exp)^2/exp))
}
```

It's best to define it early to make the work less tedious. Now...\
\
\textbf{Problem 5:} The question asks if there is a relationship between a response to a question and ethnic groups. Data will be imported here.
```{r}
#input of columns from book
obs.yes = c(78,56,43,53,43,36,42,29)
obs.no = c(47,29,29,32,30,22,23,7)
ethnic.total = obs.yes + obs.no

#this adds the last row which includes the sum of the respective col.
obs.yes = c(obs.yes, sum(obs.yes))
obs.no = c(obs.no, sum(obs.no))
ethnic.total = c(ethnic.total, sum(ethnic.total))

#following calculates the expected values for yes/no
exp.yes = tail(obs.yes,1)*ethnic.total/tail(ethnic.total, 1)
exp.no = tail(obs.no, 1)*ethnic.total/tail(ethnic.total, 1)

#calculating test statistic and p-value for each for obs/exp pair
ts1 = pears_chi(obs.yes, exp.yes) + pears_chi(obs.no, exp.no)
ts1
pchisq(ts1, df = 7, lower.tail = FALSE)
```

The null hypothesis can be stated as, any ethnicity has an equal chance of selecting yes or no. From the above test statistic and corresponding p-value, there is good evidence to suggest that ethnicity is independent from the response to the question posed.
\newpage
```{r}
#this is what the data looks like all together
data.frame(obs.yes,obs.no,ethnic.total,exp.yes,exp.no)
```

\textbf{Problem 9:} This question asks whether Austen was consistent across the works she created. Then asks if her imitator was successful in copying her style. I will first compare Sense and Sensibility with Emma.
```{r,include=FALSE}
#inputting cols and calculating cols/rows sum
obs.ss = c(14,133,12,241,11,259)
obs.ss = c(obs.ss, sum(obs.ss))
obs.emma = c(16,180,14,285,6,265)
obs.emma = c(obs.emma, sum(obs.emma))
obs.sand1 = c(8,93,12,139,8,221)
obs.sand1 = c(obs.sand1, sum(obs.sand1))
obs.sand2 = c(2,81,1,153,17,204)
obs.sand2 = c(obs.sand2, sum(obs.sand2))
```

```{r}
totalsse = obs.ss + obs.emma
exp.ss1 = tail(obs.ss, 1)*totalsse/tail(totalsse, 1)
exp.emma1 = tail(obs.emma, 1)*totalsse/tail(totalsse, 1)
ts2 = pears_chi(obs.ss,exp.ss1) + pears_chi(obs.emma,exp.emma1)
ts2
pchisq(ts2, df=5, lower.tail = FALSE)
```

The above output is the test statistic and the p-value for the homogeneity test. Both show that there is good evidence that Austen was consistent across her works Sense and Sensibility and Emma.\
\
I will now test if she was consistent across the previously tested books (pooled data) and Sanditon I.
```{r}
obs.sse = totalsse
tot2 = obs.sse + obs.sand1 #sum of pooled data and sand1

#calculate expected values for each:
exp.sse = tail(obs.sse,1)*tot2/tail(tot2,1)
exp.sand1 = tail(obs.sand1,1)*tot2/tail(tot2,1)
ts3 = pears_chi(obs.sse,exp.sse) + pears_chi(obs.sand1,exp.sand1)
ts3
pchisq(ts3, df = 5, lower.tail = FALSE)
```

The above outputs for the test statistic and the p-value suggest very strongly that Austen was not consistent across  her own books.\
\
Now I'll check whether if Sanditon I and Sandition II were consistent.
```{r}
tot.sss = obs.sand1+obs.sand2
exp.sand1 = tail(obs.sand1,1)*tot.sss/tail(tot.sss,1) #reassignment
exp.sand2 = tail(obs.sand2,1)*tot.sss/tail(tot.sss,1)
ts4 = pears_chi(obs.sand1, exp.sand1) + pears_chi(obs.sand2, exp.sand2)
ts4
pchisq(ts4,df=5,lower.tail=FALSE)
```

The above test statistic and p-value indicates that Austen's imitator was not consistent with the writing style found in Sanditon I.\
\
\textbf{Probelm 16:} The following problem includes a $3 \times 3$ table of data. The columns seem to sum up to 100, so this appears to be a test of independence.
```{r,echo=FALSE}
cautious = c(79,10,10)
obs.cautious = c(cautious, sum(cautious))
midroad = c(58,8,34)
obs.midroad = c(midroad, sum(midroad))
explorer = c(49,9,42)
obs.explorer = c(explorer, sum(explorer))
att.total = obs.cautious+obs.midroad+obs.explorer
attitudes = c('favorable','neutral','unfavorable','total')
attitudes1 = c('favorable','neutral','unfavorable')
data16 = data.matrix(data.frame(cautious,midroad,explorer))
data.frame(attitudes,obs.cautious,obs.midroad,obs.explorer,att.total)
```

```{r}
exp.cautious <- tail(obs.cautious,1)*att.total/tail(att.total,1)
exp.midroad <- tail(obs.midroad,1)*att.total/tail(att.total,1)
exp.explorer <- tail(obs.explorer,1)*att.total/tail(att.total,1)

ts5 = pears_chi(obs.cautious,exp.cautious)+pears_chi(obs.midroad,exp.midroad)+pears_chi(obs.explorer,exp.explorer)
ts5
pchisq(ts5,df=4,lower.tail = FALSE)
```

The above test statistic and corresponding p-value are strong evidence suggesting that there is a relationship between personality type and small cars. The relationship appears to go as follows: 
\begin{itemize}
\item Cautious people tend to have a more favorable attitude towards individuals driving in small cars.
\item Midroad people appear to be split between favorable and unfavorable attitudes, though more of these people appear to favor small cars. Very few people fall in the neutral ground.
\item Explorer people, are very split, with few people falling in the neutral category.
\end{itemize}
```{r,out.width='50%',out.height='50%',fig.align='center'}
rownames(data16) <- attitudes1
barplot(data16)
```

Here is the stacked bar plot of the data.\
\
\textbf{Problem 29:} This question is asking, given the scenario, would I, the company's statistician, perform a $\chi^2$ test of independence or homogeneity.\
\
The type of test to be utilized would completely depend on the way the sampling was made. For example:
\begin{itemize}
\item If the sample of 100 was made by taking random people from population of employees, then dividing them into categories of male/female, then this would require a test of independence.
\item If the amount of people to be selected from each sex were fixed, that is 50 men and 50 women, then this would require a test of homogeneity.
\end{itemize}
\
\textbf{Problem 14:} I will begin by inputting the necessary data for both of the tables in the problem:
```{r}
#first table. No high school:
under45.nohs.obs = c(71,305)
over45.nohs.obs = c(217,652)
under45.nohs.obs = c(under45.nohs.obs, sum(under45.nohs.obs))
over45.nohs.obs = c(over45.nohs.obs, sum(over45.nohs.obs))
total.nohs.obs = under45.nohs.obs + over45.nohs.obs

#second table. Some high school:
under45.some.obs = c(305,869)
over45.some.obs = c(180,259)
under45.some.obs = c(under45.some.obs, sum(under45.some.obs))
over45.some.obs = c(over45.some.obs, sum(over45.some.obs))
total.some.obs = under45.some.obs + over45.some.obs
```

\textbf{A.} This question asks to analyze the dependence of interest in political elections on age and education. The following code calculates the expected values for each of the tables, obtains a test statistics, and p-value.
```{r}
#expected values for no high school:
under45.nohs.exp = tail(under45.nohs.obs,1)*total.nohs.obs/tail(total.nohs.obs,1)
over45.nohs.exp = tail(over45.nohs.obs,1)*total.nohs.obs/tail(total.nohs.obs,1)

#expected values for some high school:
under45.some.exp = tail(under45.some.obs,1)*total.some.obs/tail(total.some.obs,1)
over45.some.exp = tail(over45.some.obs,1)*total.some.obs/tail(total.some.obs,1)

#test statistic for independence: 
ts6 = pears_chi(under45.nohs.obs,under45.nohs.exp) + pears_chi(over45.nohs.obs,over45.nohs.exp) +
  pears_chi(under45.some.obs,under45.some.exp) + pears_chi(over45.some.obs, over45.some.exp)
ts6

#p-value for independence:
pchisq(ts6, df = 2*1*1, lower.tail = FALSE)
```

The above output indicates that given interest in political elections, there appears to be strong evidence to suggest that there is a relationship between age and education.\
\
\textbf{B.} This question asks to test the following hypothesis: 
\begin{itemize}
\item $H_1:$ given education, age and interest in politics are unrelated.
\item $H_2:$ given age, education and interest are unrelated.
\end{itemize}

This first wall of code will test $H_1$.
```{r}
#observed values for under 45:
lit.und.obs = c(305,869)
great.und.obs = c(71,305) 
lit.und.obs = c(lit.und.obs, sum(lit.und.obs))
great.und.obs = c(great.und.obs, sum(great.und.obs))
tot.und = lit.und.obs + great.und.obs

#observed values for over 45:
lit.ov.obs = c(652,259)
great.ov.obs = c(217,180)
lit.ov.obs = c(lit.ov.obs, sum(lit.ov.obs))
great.ov.obs = c(great.ov.obs, sum(great.ov.obs))
tot.ov = lit.ov.obs + great.ov.obs

#expected values for under 45 table:
lit.und.exp = tail(lit.und.obs,1)*tot.und/tail(tot.und,1)
great.und.exp = tail(great.und.obs,1)*tot.und/tail(tot.und,1)

#expected values for over 45 table:
lit.ov.exp = tail(lit.ov.obs,1)*tot.ov/tail(tot.ov,1)
great.ov.exp = tail(great.ov.obs,1)*tot.ov/tail(tot.ov,1)

#test statistic for the above values:
ts7 = pears_chi(lit.und.obs, lit.und.exp) + pears_chi(great.und.obs, great.und.exp) +
  pears_chi(lit.ov.obs, lit.ov.exp) + pears_chi(great.ov.obs, great.ov.exp)
ts7

pchisq(ts7, df = 2, lower.tail = FALSE)
```

The above test statistic and p-value indicate that there is strong evidence to suggest that given education, there appears to be a relationship between age and the degree of interest in politics.

The following test statistic and p-value output will pertain to tables distinguishing between little and great interest in politics. I will refrain from outputting the code I used to prevent this assignment from being any longer.
```{r,echo=FALSE}
#great political interest:
great.some.obs = c(305,180)
great.none.obs = c(71,217)
great.some.obs = c(great.some.obs, sum(great.some.obs))
great.none.obs = c(great.none.obs, sum(great.none.obs))
tot.great = great.some.obs + great.none.obs

#little political interest:
little.some.obs = c(869,259)
little.none.obs = c(305,652)
little.some.obs = c(little.some.obs, sum(little.some.obs))
little.none.obs = c(little.none.obs, sum(little.none.obs))
tot.little = little.some.obs + little.none.obs

#expected values:
great.some.exp = tail(great.some.obs, 1)*tot.great/tail(tot.great, 1)
great.none.exp = tail(great.none.obs, 1)*tot.great/tail(tot.great, 1)
little.some.exp = tail(little.some.obs, 1)*tot.little/tail(tot.little, 1)
little.none.exp = tail(little.none.obs, 1)*tot.little/tail(tot.little, 1)

#ts
ts8 = pears_chi(great.some.obs,great.some.exp) + pears_chi(great.none.obs, great.none.exp) +
  pears_chi(little.some.obs, little.some.exp) + pears_chi(little.none.obs, little.none.exp)

ts8

#p-value
pchisq(ts8, 2, lower.tail = FALSE)
```

This test statistic is huge! It's 535.08. This alone would provide enough evidence to suggest that given age, educational level and degree of interest in politics do have a strong relationship with one another.\
\
To conclude, I rejected $H_0,\ H_1,$ and $H_2$. This means that given an one of the three parameters recorded, there is strong evidence to suggest that the other two parameters are related.